---
title: "LittorinaStatsClean"
author: "Ffion Titmuss"
date: "3/16/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

setwd("~/Repos/LittorinaThermal2019")

## Load packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(brglm2)
```

## Load data
```{r}
litt <- read.csv("./data/LittorinaData.csv") %>%
  rename(Spp = Genus_Species)

mortB23 <- litt %>%
  filter(PreTrtMortNotes == "") %>% # remove snails that died during acclimation or were misidentified
  filter(WetWeight <= 1) %>% # remove size outliers
  filter(Block == 2 | Block == 3) %>%
  mutate(Block = replace(Block, Block == 2, "A"), Block = replace(Block, Block == 3, "B")) %>%
  mutate(DaysAlive = ifelse(TrtDay2_Survived == TRUE, 2, ifelse(TrtDay1_Survived == TRUE, 1, 0))) %>%
  mutate(DaysDead = 2 - DaysAlive) %>%
  mutate(PropDaysAlive = DaysAlive/2) %>%
  mutate(Spp = as.factor(Spp)) %>%
  mutate(PopID = as.factor(PopID)) %>%
  mutate(Trt = factor(Trt, levels = c("NT", "CS", "HS"))) %>%
  mutate(Block = as.factor(Block))
```

## Final model
```{r}
mortModel <- glm(cbind(DaysAlive, DaysDead) ~ Spp * PopID * Trt * Block - Spp:PopID:Trt:Block - Spp:Trt:Block - Spp:PopID:Trt - Spp:PopID:Block - PopID:Trt:Block - PopID:Block - Spp:Block, data = mortB23, family = binomial(link = "logit"), method = "brglm_fit")

summary(mortModel)
```

## Look at model matrix
```{r}
matrix <- model.matrix(mortModel)
# Intercept: L. littorea / MA / control / block A

trtCombos <- mortB23 %>%
  group_by(Spp, PopID, Trt, Block) %>%
  summarize(Height = mean(ShellHeight)) %>%
  subset(select = -Height)

matrix2 <- model.matrix(~ Spp * PopID * Trt * Block - Spp:PopID:Trt:Block - Spp:Trt:Block - Spp:PopID:Trt - Spp:PopID:Block - PopID:Trt:Block - PopID:Block - Spp:Block, trtCombos) %>%
  as.data.frame()

matrix3 <- cbind(trtCombos, matrix2)
```

## Look at significance

### In control conditions, are there overall spp differences?
```{r}
# littorea - obtusata
# littorea - saxatilis

str(mortB23$Spp)
str(mortB23$Trt)

m2 <- brglm(cbind(DaysAlive, DaysDead) ~ Spp * PopID * Trt * Block - Spp:PopID:Trt:Block - Spp:Trt:Block - Spp:PopID:Trt - Spp:PopID:Block - PopID:Trt:Block - PopID:Block - Spp:Block, data = mortB23, family = binomial(link = "logit"), contrasts = list(Spp = contr.treatment, PopID = contr.sum, Trt = contr.treatment, Block = contr.sum))

summary(m2)

# obtusata - saxatilis

mortB23$Spp <- factor(mortB23$Spp, levels = c("littorina_obtusata", "littorina_saxatilis", "littorina_littorea"), ordered = TRUE)
str(mortB23$Spp)

m3 <- brglm(cbind(DaysAlive, DaysDead) ~ Spp * PopID * Trt * Block - Spp:PopID:Trt:Block - Spp:Trt:Block - Spp:PopID:Trt - Spp:PopID:Block - PopID:Trt:Block - PopID:Block - Spp:Block, data = mortB23, family = binomial(link = "logit"), contrasts = list(Spp = contr.treatment, PopID = contr.sum, Trt = contr.treatment, Block = contr.sum))

summary(m3)
```

### In control conditions, are there overall population differences?
```{r}
m4 <- brglm(cbind(DaysAlive, DaysDead) ~ Spp * PopID * Trt * Block - Spp:PopID:Trt:Block - Spp:Trt:Block - Spp:PopID:Trt - Spp:PopID:Block - PopID:Trt:Block - PopID:Block - Spp:Block, data = mortB23, family = binomial(link = "logit"), contrasts = list(Spp = contr.sum, PopID = contr.treatment, Trt = contr.treatment, Block = contr.sum))

summary(m4)
```

### In control conditions, what is the difference between populations within a spp?
```{r}
# littorea

mortB23$Spp <- factor(mortB23$Spp, levels = c("littorina_littorea", "littorina_obtusata", "littorina_saxatilis"), ordered = TRUE)
str(mortB23$Spp)

m5 <- brglm(cbind(DaysAlive, DaysDead) ~ Spp * PopID * Trt * Block - Spp:PopID:Trt:Block - Spp:Trt:Block - Spp:PopID:Trt - Spp:PopID:Block - PopID:Trt:Block - PopID:Block - Spp:Block, data = mortB23, family = binomial(link = "logit"), contrasts = list(Spp = contr.treatment, PopID = contr.treatment, Trt = contr.treatment, Block = contr.sum))

summary(m5)

# obtusata

mortB23$Spp <- factor(mortB23$Spp, levels = c("littorina_obtusata", "littorina_saxatilis", "littorina_littorea"), ordered = TRUE)
str(mortB23$Spp)

m6 <- brglm(cbind(DaysAlive, DaysDead) ~ Spp * PopID * Trt * Block - Spp:PopID:Trt:Block - Spp:Trt:Block - Spp:PopID:Trt - Spp:PopID:Block - PopID:Trt:Block - PopID:Block - Spp:Block, data = mortB23, family = binomial(link = "logit"), contrasts = list(Spp = contr.treatment, PopID = contr.treatment, Trt = contr.treatment, Block = contr.sum))

summary(m6)

# saxatilis

mortB23$Spp <- factor(mortB23$Spp, levels = c("littorina_saxatilis", "littorina_littorea", "littorina_obtusata"), ordered = TRUE)
str(mortB23$Spp)

m7 <- brglm(cbind(DaysAlive, DaysDead) ~ Spp * PopID * Trt * Block - Spp:PopID:Trt:Block - Spp:Trt:Block - Spp:PopID:Trt - Spp:PopID:Block - PopID:Trt:Block - PopID:Block - Spp:Block, data = mortB23, family = binomial(link = "logit"), contrasts = list(Spp = contr.treatment, PopID = contr.treatment, Trt = contr.treatment, Block = contr.sum))

summary(m7)
```

### In heat conditions, are there overall spp differences?
```{r}
# littorea - obtusata
# littorea - saxatilis

mortB23$Spp <- factor(mortB23$Spp, levels = c("littorina_littorea", "littorina_obtusata", "littorina_saxatilis"), ordered = TRUE)
str(mortB23$Spp)

mortB23$Trt <- factor(mortB23$Trt, levels = c("HS", "CS", "NT"), ordered = TRUE)
str(mortB23$Trt)

m8 <- brglm(cbind(DaysAlive, DaysDead) ~ Spp * PopID * Trt * Block - Spp:PopID:Trt:Block - Spp:Trt:Block - Spp:PopID:Trt - Spp:PopID:Block - PopID:Trt:Block - PopID:Block - Spp:Block, data = mortB23, family = binomial(link = "logit"), contrasts = list(Spp = contr.treatment, PopID = contr.sum, Trt = contr.treatment, Block = contr.sum))

summary(m8)

# obtusata - saxatilis

mortB23$Spp <- factor(mortB23$Spp, levels = c("littorina_obtusata", "littorina_saxatilis", "littorina_littorea"), ordered = TRUE)
str(mortB23$Spp)

m9 <- brglm(cbind(DaysAlive, DaysDead) ~ Spp * PopID * Trt * Block - Spp:PopID:Trt:Block - Spp:Trt:Block - Spp:PopID:Trt - Spp:PopID:Block - PopID:Trt:Block - PopID:Block - Spp:Block, data = mortB23, family = binomial(link = "logit"), contrasts = list(Spp = contr.treatment, PopID = contr.sum, Trt = contr.treatment, Block = contr.sum))

summary(m9)
```

### In heat conditions, are there overall population differences?
```{r}
m10 <- brglm(cbind(DaysAlive, DaysDead) ~ Spp * PopID * Trt * Block - Spp:PopID:Trt:Block - Spp:Trt:Block - Spp:PopID:Trt - Spp:PopID:Block - PopID:Trt:Block - PopID:Block - Spp:Block, data = mortB23, family = binomial(link = "logit"), contrasts = list(Spp = contr.sum, PopID = contr.treatment, Trt = contr.treatment, Block = contr.sum))

summary(m10)
```

### In heat conditions, what is the difference between populations within a spp?
```{r}
# littorea

mortB23$Spp <- factor(mortB23$Spp, levels = c("littorina_littorea", "littorina_obtusata", "littorina_saxatilis"), ordered = TRUE)
str(mortB23$Spp)

m11 <- brglm(cbind(DaysAlive, DaysDead) ~ Spp * PopID * Trt * Block - Spp:PopID:Trt:Block - Spp:Trt:Block - Spp:PopID:Trt - Spp:PopID:Block - PopID:Trt:Block - PopID:Block - Spp:Block, data = mortB23, family = binomial(link = "logit"), contrasts = list(Spp = contr.treatment, PopID = contr.treatment, Trt = contr.treatment, Block = contr.sum))

summary(m11)

# obtusata

mortB23$Spp <- factor(mortB23$Spp, levels = c("littorina_obtusata", "littorina_saxatilis", "littorina_littorea"), ordered = TRUE)
str(mortB23$Spp)

m12 <- brglm(cbind(DaysAlive, DaysDead) ~ Spp * PopID * Trt * Block - Spp:PopID:Trt:Block - Spp:Trt:Block - Spp:PopID:Trt - Spp:PopID:Block - PopID:Trt:Block - PopID:Block - Spp:Block, data = mortB23, family = binomial(link = "logit"), contrasts = list(Spp = contr.treatment, PopID = contr.treatment, Trt = contr.treatment, Block = contr.sum))

summary(m12)

# saxatilis

mortB23$Spp <- factor(mortB23$Spp, levels = c("littorina_saxatilis", "littorina_littorea", "littorina_obtusata"), ordered = TRUE)
str(mortB23$Spp)

m13 <- brglm(cbind(DaysAlive, DaysDead) ~ Spp * PopID * Trt * Block - Spp:PopID:Trt:Block - Spp:Trt:Block - Spp:PopID:Trt - Spp:PopID:Block - PopID:Trt:Block - PopID:Block - Spp:Block, data = mortB23, family = binomial(link = "logit"), contrasts = list(Spp = contr.treatment, PopID = contr.treatment, Trt = contr.treatment, Block = contr.sum))

summary(m13)
```

### In cold conditions, are there overall spp differences?
```{r}
# littorea - obtusata
# littorea - saxatilis

mortB23$Spp <- factor(mortB23$Spp, levels = c("littorina_littorea", "littorina_obtusata", "littorina_saxatilis"), ordered = TRUE)
str(mortB23$Spp)

mortB23$Trt <- factor(mortB23$Trt, levels = c("CS", "NT", "HS"), ordered = TRUE)
str(mortB23$Trt)

m14 <- brglm(cbind(DaysAlive, DaysDead) ~ Spp * PopID * Trt * Block - Spp:PopID:Trt:Block - Spp:Trt:Block - Spp:PopID:Trt - Spp:PopID:Block - PopID:Trt:Block - PopID:Block - Spp:Block, data = mortB23, family = binomial(link = "logit"), contrasts = list(Spp = contr.treatment, PopID = contr.sum, Trt = contr.treatment, Block = contr.sum))

summary(m14)

# obtusata - saxatilis

mortB23$Spp <- factor(mortB23$Spp, levels = c("littorina_obtusata", "littorina_saxatilis", "littorina_littorea"), ordered = TRUE)
str(mortB23$Spp)

m15 <- brglm(cbind(DaysAlive, DaysDead) ~ Spp * PopID * Trt * Block - Spp:PopID:Trt:Block - Spp:Trt:Block - Spp:PopID:Trt - Spp:PopID:Block - PopID:Trt:Block - PopID:Block - Spp:Block, data = mortB23, family = binomial(link = "logit"), contrasts = list(Spp = contr.treatment, PopID = contr.sum, Trt = contr.treatment, Block = contr.sum))

summary(m15)
```

### In cold conditions, are there overall population differences?
```{r}
m16 <- brglm(cbind(DaysAlive, DaysDead) ~ Spp * PopID * Trt * Block - Spp:PopID:Trt:Block - Spp:Trt:Block - Spp:PopID:Trt - Spp:PopID:Block - PopID:Trt:Block - PopID:Block - Spp:Block, data = mortB23, family = binomial(link = "logit"), contrasts = list(Spp = contr.sum, PopID = contr.treatment, Trt = contr.treatment, Block = contr.sum))

summary(m16)
```

### In cold conditions, what is the difference between populations within a spp?
```{r}
# littorea

mortB23$Spp <- factor(mortB23$Spp, levels = c("littorina_littorea", "littorina_obtusata", "littorina_saxatilis"), ordered = TRUE)
str(mortB23$Spp)

m17 <- brglm(cbind(DaysAlive, DaysDead) ~ Spp * PopID * Trt * Block - Spp:PopID:Trt:Block - Spp:Trt:Block - Spp:PopID:Trt - Spp:PopID:Block - PopID:Trt:Block - PopID:Block - Spp:Block, data = mortB23, family = binomial(link = "logit"), contrasts = list(Spp = contr.treatment, PopID = contr.treatment, Trt = contr.treatment, Block = contr.sum))

summary(m17)

# obtusata

mortB23$Spp <- factor(mortB23$Spp, levels = c("littorina_obtusata", "littorina_saxatilis", "littorina_littorea"), ordered = TRUE)
str(mortB23$Spp)

m18 <- brglm(cbind(DaysAlive, DaysDead) ~ Spp * PopID * Trt * Block - Spp:PopID:Trt:Block - Spp:Trt:Block - Spp:PopID:Trt - Spp:PopID:Block - PopID:Trt:Block - PopID:Block - Spp:Block, data = mortB23, family = binomial(link = "logit"), contrasts = list(Spp = contr.treatment, PopID = contr.treatment, Trt = contr.treatment, Block = contr.sum))

summary(m18)

# saxatilis

mortB23$Spp <- factor(mortB23$Spp, levels = c("littorina_saxatilis", "littorina_littorea", "littorina_obtusata"), ordered = TRUE)
str(mortB23$Spp)

m19 <- brglm(cbind(DaysAlive, DaysDead) ~ Spp * PopID * Trt * Block - Spp:PopID:Trt:Block - Spp:Trt:Block - Spp:PopID:Trt - Spp:PopID:Block - PopID:Trt:Block - PopID:Block - Spp:Block, data = mortB23, family = binomial(link = "logit"), contrasts = list(Spp = contr.treatment, PopID = contr.treatment, Trt = contr.treatment, Block = contr.sum))

summary(m19)
```

### Alternate method

litt_HS <- mortB23 %>%
  filter(Spp == "littorina_littorea") %>%
  filter(Trt == "HS")

m11.1 <- brglm(cbind(DaysAlive, DaysDead) ~ PopID * Block, data = litt_HS, family = binomial(link = "logit"), contrasts = list(PopID = contr.treatment, Block = contr.sum))

summary(m11.1)

CS <- mortB23 %>%
  filter(Trt == "CS")

m16.1 <- brglm(cbind(DaysAlive, DaysDead) ~ Spp * PopID * Block, data = CS, family = binomial(link = "logit"), contrasts = list(Spp = contr.sum, PopID = contr.treatment, Block = contr.sum))

summary(m16.1)

litt_CS <- mortB23 %>%
  filter(Spp == "littorina_littorea") %>%
  filter(Trt == "CS")

m17.1 <- brglm(cbind(DaysAlive, DaysDead) ~ PopID * Block, data = litt_CS, family = binomial(link = "logit"), contrasts = list(PopID = contr.treatment, Block = contr.sum))

summary(m17.1)

obt_CS <- mortB23 %>%
  filter(Spp == "littorina_obtusata") %>%
  filter(Trt == "CS")

m18.1 <- brglm(cbind(DaysAlive, DaysDead) ~ PopID * Block, data = obt_CS, family = binomial(link = "logit"), contrasts = list(PopID = contr.treatment, Block = contr.sum))

summary(m18.1)