---
title: "Littorina CovGE"
author: "Molly Albecker"
date: "4/8/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

setwd("~/Repos/LittorinaThermal2019")

Here we provide an example for the covariance analysis. We use the dataset on amphibian (frog) development that was provided by Albecker and McCoy 2019 and is an example presented in the paper. 

First steps are to load the data and the functions. These are available on the Github (https://github.com/RCN-ECS/CnGV/tree/master/src/)
```{r}

# Packages
if (!require('lme4')) install.packages('lme4'); library('lme4')
if (!require('emmeans')) install.packages('emmeans'); library('emmeans')
if (!require('tidyverse')) install.packages('tidyverse'); library('tidyverse')

# Functions (may need to change your working directory)
source("./src/CovarianceDataFunctions.R")

# Overall Dataset
snail1 = read.csv("./data/LittorinaData.csv")

```

Let's format the data! I'm going to separate these out into different datasets for each species. I will also add/rename columns just so they work in the functions. 

gen_factor = population (genotype), 
exp_env_factor = experimental treatment, 
nat_env_factor = native environment

We also need to ensure the data are formatted for the analysis. We use an ANOVA to extract estimated marginal means for each genotype and environmental mean (yi and yj in CovGE equation), which requires that fixed effects (genotype and environment) are categorical. 

Finally, we'll analyze each phenotype and species differently.

```{r}

# Exclude block 1 and control treatments (only want data from environments that have a corresponding "genotype" )
snail1$gen_factor = snail1$exp_env_factor = snail1$nat_env_factor = snail1$phen_data = NULL
snail = snail1 %>% 
  filter(Block != 1) %>% 
  filter(Trt != "NT") %>%
  mutate(nat_env_factor = ifelse(PopID == "MA", "Cold", "Warm")) %>%
  #rename(gen_factor = PopID) %>%
  mutate(exp_env_factor = ifelse(Trt == "CS", "Cold", "Warm")) %>%
  mutate(nat_env_factor = as.factor(nat_env_factor))%>%
  mutate(gen_factor = ifelse(PopID == "MA", "Cold", "Warm"))%>%
  mutate(exp_env_factor = as.factor(exp_env_factor)) %>%
  mutate(gen_factor = as.factor(gen_factor)) %>%
  mutate(DaysAlive = ifelse(TrtDay2_Survived == TRUE, 2, ifelse(TrtDay1_Survived == TRUE, 1, 0))) %>%
  mutate(DaysDead = 2 - DaysAlive) %>%
  mutate(PropDaysAlive = DaysAlive/2) %>%
  rename(phen_data = PropDaysAlive) %>%
  filter(Genus_Species == "littorina_saxatilis")  ### Here is where you sub in "littorina_saxatilis", "littorina_obtusata", or "littorina_littorea" to run the three species through

```


##### Standardize data: 

We standardize the phenotypic data by subtracting the overall mean phenotype from each phenotypic datapoint and dividing by the standard deviation of group means (group = genotype and experimental environment pair).

```{r}
 
snail$group = paste(snail$gen_factor,snail$exp_env_factor,sep = "-")
snail$phen_corrected = (snail$phen_data - mean(snail$phen_data, na.rm = TRUE))/sd(tapply(snail$phen_data, snail$group, mean, na.rm = TRUE))

```

## Calculate CovGE: 

To start out, we ran a basic categorical linear model to generate yi and yj parameters via estimated marginal means. Estimated marginal means are more robust given unbalanced study designs.   

After running the model, we extract estimated marginal means using function emmeans(). 

```{r}
  
# Anova
aov.test <- lm(phen_corrected ~ exp_env_factor * gen_factor, data = snail) 

# Estimated Marginal Means
emm_df = as.data.frame(emmeans(aov.test, ~ exp_env_factor*gen_factor))

```
This gives us the matrix of yij that will be used for GxE. But for CovGE, we next we need to calculate yi (genotypic means) and yj (experimental environment means). 

We do this using tapply() to calculate the mean phenotype for each genotype ACROSS environments (G_matrix), and then the mean phenotype for each environment ACROSS genotypes (E_matrix).

Because this is a common garden design, there should be more genotypic means than environmental means.
```{r}

G_matrix <- data.frame("G_means" = tapply(emm_df$emmean, emm_df$gen_factor, mean, na.rm=TRUE),
                       "gen_factor" = unique(emm_df$gen_factor))

E_matrix <- data.frame("E_means" = tapply(emm_df$emmean, emm_df$exp_env_factor, mean, na.rm=TRUE),
                       "exp_env_factor" = unique(emm_df$exp_env_factor))

```
 
To match each genotypic mean with the correct environmental mean, we have to ensure genotypes are correctly matched to their native environment. This is what the "I" term in the CovGE equation refers to. 

Because there are more genotypes than environments, environments will be used more than once. 
```{r} 
# First create Native Environment reference dataframe 
native_df = data.frame("gen_factor" = unique(snail$gen_factor))
native_df$nat_env_factor = snail$nat_env_factor[match(native_df$gen_factor, snail$gen_factor)]

# Reorder Gmatrix and Ematrix to reflect the above native environments 
Cov_matrix = G_matrix
Cov_matrix$exp_env_factor = native_df$nat_env_factor[match(G_matrix$gen_factor,native_df$gen_factor)] # Indicator Variable
Cov_matrix$E_means = E_matrix$E_means[match(Cov_matrix$exp_env_factor,E_matrix$exp_env_factor)]

```

Now we have the data formatted so that we can calculate CovGE. 

```{r}
N = length(Cov_matrix$gen_factor) # Length of number of genotypes
overallmean = mean(emm_df$emmean,na.rm=TRUE) # ybar (overall mean phenotype across all genotypes and environments)
numerator = sum((Cov_matrix$G_means - overallmean)*(Cov_matrix$E_means - overallmean)) # Follows Numerator in CovGE. 
  
standardize_max = max(var(Cov_matrix$E_means),var(Cov_matrix$G_means)) # standardize CovGE by max variance of either Gmeans or Emeans
CovGE = (1/(N-1))*(numerator/standardize_max) 
CovGE
```

Woah - for Littorina saxatilis - I'm getting a reading of -0.9954! THAT'S REALLY STRONG COUNTERGRADIENT VARIATION! Let's get confidence intervals and P-values next.


## Bootstrapped Confidence Intervals

To estimate confidence intervals, we use bootstrapping, in which we shuffle phenotype within each genotype/environment and recalculate covGE after each reshuffle. This generates a distribution of CovGE estimates that form 95% confidence intervals. 

For this step, I am going to use imported functions that do the same thing as shown above.
"bootstrap_raw" function shuffles the raw datadata,
"mod.GxE" function compiles the cov_matrix dataframe
"cov.function" function calculates CovGE

```{r}
n_boot <- 999
balanced = TRUE   # This would be in case your data were unbalanced (different numbers of genotypes)

boot_dat_raw = boot_df_raw = data.frame()
    
for(i in 1:n_boot){
      
  # Shuffle Data
  shuffle_dat <- bootstrap_raw(snail) 
  
  # Anova model fit & GxE estimates
  m2 <- mod.Cov(shuffle_dat,balanced) # Insert shuffled raw phenotype dataframe
  
  # Pull info from mod.Cov output
  cov_matrix_boot <- m2[[1]]
  
  # Covariance Estimates
  cov_corrected_boot = round(cov.function(cov_matrix_boot, balanced),3)
  
  # Bootstrap dataframe
  boot_dat_raw <- data.frame("covariance_boot" = cov_corrected_boot)
  boot_df_raw <- rbind(boot_df_raw,boot_dat_raw)
}
    
# Check: Histograms of distribution around CovGE (sanity check)
ggplot(boot_df_raw, aes(x = covariance_boot), alpha = 0.5)+  geom_histogram()+ geom_vline(aes(xintercept = CovGE))+
  ggtitle("Bootstrapped CovGE values")
    
    
# Covariance Confidence Intervals 
cov_CI = quantile(boot_df_raw$covariance_boot, probs=c(0.025, 0.975), type=1) 
cov_CI    

```
Alright!  95% Confidence intervals! Let's see if it's significant using permutation. 


## Hypothesis Testing using Permutation

For hypothesis testing, we used permutation. Permutation also resamples phenotypic data but does not resample or maintain the genotypic/environmental levels. As a result, it creates a distribution around the null expectation that CovGE = 0. If the CovGE estimate (-0.42) is outside of the tails of this null distribution, it is considered statistically significant. 

Again, I will used some canned functions that simply speed up the above CovGE calculation.
"permutation_raw" permutes data.
"mod.Cov" again takes those permuted data and generates Cov_matrix dataframes
"cov.function" calculates CovGE of permuted data


```{r}

# Output dataframe
perm_df_raw = perm_dat_raw = data.frame()
    
for(i in 1:n_boot){
  
  # Resample Data
  perm_dat <- permutation_raw(snail)
  
  # Anova model fit & GxE estimates
  m3 <- mod.Cov(perm_dat, balanced) # Insert permuted data
  
  # GxE Estimates
  cov_matrix_perm <- m3[[1]]

  # Covariance Estimates
  cov_corrected_perm = round(cov.function(cov_matrix_perm, balanced),3)
  
  # Permutation dataframe
  perm_dat_raw <- data.frame("covariance_perm" = cov_corrected_perm)
  perm_df_raw <- rbind(perm_df_raw,perm_dat_raw)
}

# Check: Permutation histogram - should be around zero
ggplot(perm_df_raw, aes(x = covariance_perm), alpha = 0.5)+ geom_histogram()  + ggtitle("Null Distribution for CovGE") # Should be centered around 0
```

Now to calculate the p-value from this null distribution: 
```{r}
# Covariance P-values
cov_pvalue <- sum(abs(perm_df_raw$covariance_perm) >= abs(CovGE))/(n_boot+1) # Two-tailed
cov_pvalue     
```

Congratulations! You have estimated CovGE, 95% confidence intervals, and the P-value for your snail data.
